Meta-analytic approaches and 'new' effect sizes to account for heterogeneous effects in comparative physiology

### 2021-07-08 MS/ms.Rmd
Added sampling variance eqn from Nakagawa and Santos.

where $\sigma^2_{total} = \sigma^2_{study} + \sigma^2_{phylogeny} + \sigma^2_{species} + \sigma^2_{residual} +\sigma^2_{m}$ is the total effect size variance  and $\sigma^2_{m}$ is the sampling error variance calculated as:
$$

$$
Quantifying heterogeneity using $I_{total}^2$ provides a formal way to assess the relative amount of variation that is the result of 'true' biological or methodological differences as opposed to simply chance (i.e., sampling variance) [@NakagawaSantos2012; @Higgins2002]. In other words, $I_{total}^2$ can be thought of as the percentage of variance between effect sizes that cannot be attributable to sampling error [@Higgins2002]. We can only calculate this metric if we have an estimate of the sampling variance for a given effect measure because only then can we quantify the proportion of variation that is the result of sampling variability. @Senior2016 have shown that total heterogeneity is often extremely high in ecological and evolutionary meta-analysis (\~91%), suggesting that effect measures vary widely across and within studies. Some of the variation is clearly the result of working with different species or strains, different traits and different outcome measures, however, some of this variation also arises because of methodological differences across studies that might be related to the dosages or temperatures used for treatments. Formally quantifying and presenting heterogeneity estimates from meta-analytic models can provide a way to understand the major drivers of effect size variation by allowing researchers to produce various $I^2$ estimates which can be formally compared within and across studies. For example, we could fit the multilevel model described above to understand what proportion of variation is the result of, say, 'study' effects following @NakagawaSantos2012: $I^2_{study} = \frac{\sigma^2_{study}}{\sigma^2_{total}}$ where $\sigma^2_{study}$ is the study specific variance, and all other variance components are for phylogeny, species, residual (effect size level) and m (total measurement/sampling error). Many readers will notice the similarities between $I^2$ and $R^2$ [@Nakgawa2013; @Nakagawa2010]. If we want to formally assess how much effect size variation is the result of methodological factors, such as temperature and dosage differences applied across studies, then we could fit our multilevel meta-regression model, including dosage or temperature (linear, quadratic or even more complex fits), and estimate how much variation is explained by these fixed effects following @Nakgawa2013.

>>>>

where $\sigma^2_{total} = \sigma^2_{study} + \sigma^2_{phylogeny} + \sigma^2_{species} + \sigma^2_{residual} +\sigma^2_{m}$ is the total effect size variance  and $\sigma^2_{m}$ is the sampling error variance calculated as:

$$
\sigma_{m}^2 = \sum w_{j}\left( k-1\right) / \left[ \left( \sum w_{j}\right)^2  + \sum w_{j}^2\right]
$$

where *k* is the number of studies and the weights, $w_{j} = \frac{1}{v_{j}}$, can be calcualted as the inverse of the sampling variance ($v_{j}$) for each effect size, *j*.

Quantifying heterogeneity using $I_{total}^2$ provides a formal way to assess the relative amount of variation that is the result of 'true' biological or methodological differences as opposed to simply chance (i.e., sampling variance) [@NakagawaSantos2012; @Higgins2002]. In other words, $I_{total}^2$ can be thought of as the percentage of variance between effect sizes that cannot be attributable to sampling error [@Higgins2002]. We can only calculate this metric if we have an estimate of the sampling variance for a given effect measure because only then can we quantify the proportion of variation that is the result of sampling variability. @Senior2016 have shown that total heterogeneity is often extremely high in ecological and evolutionary meta-analysis (\~91%), suggesting that effect measures vary widely across and within studies. Some of the variation is clearly the result of working with different species or strains, different traits and different outcome measures, however, some of this variation also arises because of methodological differences across studies that might be related to the dosages or temperatures used for treatments. Formally quantifying and presenting heterogeneity estimates from meta-analytic models can provide a way to understand the major drivers of effect size variation by allowing researchers to produce various $I^2$ estimates which can be formally compared within and across studies. For example, we could fit the multilevel model described above to understand what proportion of variation is the result of, say, 'study' effects following @NakagawaSantos2012: $I^2_{study} = \frac{\sigma^2_{study}}{\sigma^2_{total}}$ where $\sigma^2_{study}$ is the study specific variance.

Many readers will notice the similarities between $I^2$ and $R^2$ [@Nakgawa2013; @Nakagawa2010]. If we want to formally assess how much effect size variation is the result of methodological factors, such as temperature and dosage differences applied across studies, then we could fit our multilevel meta-regression model, including dosage or temperature (linear, quadratic or even more complex fits), and estimate how much variation is explained by these fixed effects following @Nakgawa2013.
